---
title: "Hyde 3.3 Athromes Processing (Figs 1C and Fig 2)"
author: "JAH"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
---

```{r include=FALSE}
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE, echo = TRUE, message = TRUE, warning = TRUE)
```

# Background:

The file containes analyses for Figures 1C and Figure 2 in the manuscript.  The HYDE 3.3 
(https://doi.org/10.24416/UU01-94FNH0) dataset is read in and processed to 3-class Anthromes (wildlands vs. cultured vs. intensive land areas)


# Fig 1C

```{r}
### make a plot 
library(ggplot2)
```


## read in ascii data
```{r}
anth_1900 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1900AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1910 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1910AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1920 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1920AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1930 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1930AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1940 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1940AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1950 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1950AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")
```

```{r}
anth_1951 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1951AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1952 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1952AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1953 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1953AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1954 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1954AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1955 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1955AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1956 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1956AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1957 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1957AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1958 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1958AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1959 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1959AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1960 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1960AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")
```

```{r}
anth_1961 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1961AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1962 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1962AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1963 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1963AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1964 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1964AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1965 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1965AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1966 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1966AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1967 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1967AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1968 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1968AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1969 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1969AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1970 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1970AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

```

```{r}
anth_1971 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1971AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1972 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1972AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1973 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1973AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1974 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1974AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1975 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1975AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1976 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1976AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1977 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1977AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1978 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1978AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1979 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1979AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1980 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1980AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

```



```{r}
anth_1981 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1981AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1982 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1982AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1983 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1983AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1984 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1984AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1985 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1985AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1986 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1986AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1987 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1987AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1988 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1988AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1989 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1989AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1990 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1990AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

```

```{r}
anth_1991 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1991AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1992 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1992AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1993 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1993AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1994 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1994AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1995 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1995AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1996 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1996AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1997 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1997AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1998 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1998AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_1999 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes1999AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2000 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2000AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

```

```{r}
anth_2001 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2001AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2002 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2002AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2003 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2003AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2004 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2004AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2005 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2005AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2006 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2006AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2007 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2007AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2008 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2008AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2009 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2009AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2010 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2010AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

```


```{r}
anth_2011 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2011AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2012 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2012AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2013 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2013AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2014 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2014AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2015 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2015AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2016 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2016AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2017 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2017AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2018 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2018AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2019 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2019AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2020 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2020AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2021 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2021AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2022 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2022AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

anth_2023 <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/anthromes2023AD.asc", sep = " ", header = FALSE, skip = 6,  na.strings = "-1")

```



```{r}
g_area <- read.table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/general_files/garea_cr.asc", sep = " ", header = FALSE, skip = 6, na.strings = "-9999")

g_area <- g_area[,1:4320]
```


```{r}
### make a massive dataset
HYDE_anthromes <-  as.data.frame(cbind(unlist(anth_1900), unlist(anth_1910), unlist(anth_1920), 
                                       unlist(anth_1930), unlist(anth_1940), unlist(anth_1950), 
                        unlist(anth_1951), unlist(anth_1952), unlist(anth_1953), unlist(anth_1954), unlist(anth_1955),
                        unlist(anth_1956), unlist(anth_1957), unlist(anth_1958), unlist(anth_1959), unlist(anth_1960), 
                        unlist(anth_1961), unlist(anth_1962), unlist(anth_1963), unlist(anth_1964), unlist(anth_1965),
                        unlist(anth_1966), unlist(anth_1967), unlist(anth_1968), unlist(anth_1969), unlist(anth_1970), 
                        unlist(anth_1971), unlist(anth_1972), unlist(anth_1973), unlist(anth_1974), unlist(anth_1975),
                        unlist(anth_1976), unlist(anth_1977), unlist(anth_1978), unlist(anth_1979), unlist(anth_1980), 
                        unlist(anth_1981), unlist(anth_1982), unlist(anth_1983), unlist(anth_1984), unlist(anth_1985),
                        unlist(anth_1986), unlist(anth_1987), unlist(anth_1988), unlist(anth_1989), unlist(anth_1990), 
                        unlist(anth_1991), unlist(anth_1992), unlist(anth_1993), unlist(anth_1994), unlist(anth_1995),
                        unlist(anth_1996), unlist(anth_1997), unlist(anth_1998), unlist(anth_1999), unlist(anth_2000), 
                        unlist(anth_2001), unlist(anth_2002), unlist(anth_2003), unlist(anth_2004), unlist(anth_2005),
                        unlist(anth_2006), unlist(anth_2007), unlist(anth_2008), unlist(anth_2009), unlist(anth_2010), 
                        unlist(anth_2011), unlist(anth_2012), unlist(anth_2013), unlist(anth_2014), unlist(anth_2015),
                        unlist(anth_2016), unlist(anth_2017), unlist(anth_2018), unlist(anth_2019), unlist(anth_2020), 
                        unlist(anth_2021), unlist(anth_2022), unlist(anth_2023), unlist(g_area)))


colnames(HYDE_anthromes) <- c(seq(1900, 1950, by = 10), seq(1951, 2023, by = 1), "area")

HYDE_anthromes$`1900` <- as.factor(HYDE_anthromes$`1900`)
HYDE_anthromes$`1910` <- as.factor(HYDE_anthromes$`1910`)
HYDE_anthromes$`1920` <- as.factor(HYDE_anthromes$`1920`)
HYDE_anthromes$`1930` <- as.factor(HYDE_anthromes$`1930`)
HYDE_anthromes$`1940` <- as.factor(HYDE_anthromes$`1940`)
HYDE_anthromes$`1950` <- as.factor(HYDE_anthromes$`1950`)

HYDE_anthromes$`1951` <- as.factor(HYDE_anthromes$`1951`)
HYDE_anthromes$`1952` <- as.factor(HYDE_anthromes$`1952`)
HYDE_anthromes$`1954` <- as.factor(HYDE_anthromes$`1954`)
HYDE_anthromes$`1955` <- as.factor(HYDE_anthromes$`1955`)
HYDE_anthromes$`1956` <- as.factor(HYDE_anthromes$`1956`)
HYDE_anthromes$`1957` <- as.factor(HYDE_anthromes$`1957`)
HYDE_anthromes$`1958` <- as.factor(HYDE_anthromes$`1958`)
HYDE_anthromes$`1959` <- as.factor(HYDE_anthromes$`1959`)
HYDE_anthromes$`1960` <- as.factor(HYDE_anthromes$`1960`)

HYDE_anthromes$`1961` <- as.factor(HYDE_anthromes$`1961`)
HYDE_anthromes$`1962` <- as.factor(HYDE_anthromes$`1962`)
HYDE_anthromes$`1964` <- as.factor(HYDE_anthromes$`1964`)
HYDE_anthromes$`1965` <- as.factor(HYDE_anthromes$`1965`)
HYDE_anthromes$`1966` <- as.factor(HYDE_anthromes$`1966`)
HYDE_anthromes$`1967` <- as.factor(HYDE_anthromes$`1967`)
HYDE_anthromes$`1968` <- as.factor(HYDE_anthromes$`1968`)
HYDE_anthromes$`1969` <- as.factor(HYDE_anthromes$`1969`)
HYDE_anthromes$`1970` <- as.factor(HYDE_anthromes$`1970`)

HYDE_anthromes$`1971` <- as.factor(HYDE_anthromes$`1971`)
HYDE_anthromes$`1972` <- as.factor(HYDE_anthromes$`1972`)
HYDE_anthromes$`1974` <- as.factor(HYDE_anthromes$`1974`)
HYDE_anthromes$`1975` <- as.factor(HYDE_anthromes$`1975`)
HYDE_anthromes$`1976` <- as.factor(HYDE_anthromes$`1976`)
HYDE_anthromes$`1977` <- as.factor(HYDE_anthromes$`1977`)
HYDE_anthromes$`1978` <- as.factor(HYDE_anthromes$`1978`)
HYDE_anthromes$`1979` <- as.factor(HYDE_anthromes$`1979`)
HYDE_anthromes$`1980` <- as.factor(HYDE_anthromes$`1980`)

HYDE_anthromes$`1981` <- as.factor(HYDE_anthromes$`1981`)
HYDE_anthromes$`1982` <- as.factor(HYDE_anthromes$`1982`)
HYDE_anthromes$`1984` <- as.factor(HYDE_anthromes$`1984`)
HYDE_anthromes$`1985` <- as.factor(HYDE_anthromes$`1985`)
HYDE_anthromes$`1986` <- as.factor(HYDE_anthromes$`1986`)
HYDE_anthromes$`1987` <- as.factor(HYDE_anthromes$`1987`)
HYDE_anthromes$`1988` <- as.factor(HYDE_anthromes$`1988`)
HYDE_anthromes$`1989` <- as.factor(HYDE_anthromes$`1989`)
HYDE_anthromes$`1990` <- as.factor(HYDE_anthromes$`1990`)

HYDE_anthromes$`1991` <- as.factor(HYDE_anthromes$`1991`)
HYDE_anthromes$`1992` <- as.factor(HYDE_anthromes$`1992`)
HYDE_anthromes$`1994` <- as.factor(HYDE_anthromes$`1994`)
HYDE_anthromes$`1995` <- as.factor(HYDE_anthromes$`1995`)
HYDE_anthromes$`1996` <- as.factor(HYDE_anthromes$`1996`)
HYDE_anthromes$`1997` <- as.factor(HYDE_anthromes$`1997`)
HYDE_anthromes$`1998` <- as.factor(HYDE_anthromes$`1998`)
HYDE_anthromes$`1999` <- as.factor(HYDE_anthromes$`1999`)
HYDE_anthromes$`2000` <- as.factor(HYDE_anthromes$`2000`)

HYDE_anthromes$`2001` <- as.factor(HYDE_anthromes$`2001`)
HYDE_anthromes$`2002` <- as.factor(HYDE_anthromes$`2002`)
HYDE_anthromes$`2004` <- as.factor(HYDE_anthromes$`2004`)
HYDE_anthromes$`2005` <- as.factor(HYDE_anthromes$`2005`)
HYDE_anthromes$`2006` <- as.factor(HYDE_anthromes$`2006`)
HYDE_anthromes$`2007` <- as.factor(HYDE_anthromes$`2007`)
HYDE_anthromes$`2008` <- as.factor(HYDE_anthromes$`2008`)
HYDE_anthromes$`2009` <- as.factor(HYDE_anthromes$`2009`)
HYDE_anthromes$`2010` <- as.factor(HYDE_anthromes$`2010`)

HYDE_anthromes$`2011` <- as.factor(HYDE_anthromes$`2011`)
HYDE_anthromes$`2012` <- as.factor(HYDE_anthromes$`2012`)
HYDE_anthromes$`2014` <- as.factor(HYDE_anthromes$`2014`)
HYDE_anthromes$`2015` <- as.factor(HYDE_anthromes$`2015`)
HYDE_anthromes$`2016` <- as.factor(HYDE_anthromes$`2016`)
HYDE_anthromes$`2017` <- as.factor(HYDE_anthromes$`2017`)
HYDE_anthromes$`2018` <- as.factor(HYDE_anthromes$`2018`)
HYDE_anthromes$`2019` <- as.factor(HYDE_anthromes$`2019`)
HYDE_anthromes$`2020` <- as.factor(HYDE_anthromes$`2020`)
HYDE_anthromes$`2021` <- as.factor(HYDE_anthromes$`2021`)
HYDE_anthromes$`2022` <- as.factor(HYDE_anthromes$`2022`)
HYDE_anthromes$`2023` <- as.factor(HYDE_anthromes$`2023`)


```

```{r}
## remove other things--- clean workspace
rm(anth_1900,anth_1910,anth_1920,anth_1930,anth_1940,anth_1950, 
   anth_1951,anth_1952,anth_1953,anth_1954,anth_1955,
   anth_1956,anth_1957,anth_1958,anth_1959,anth_1960, 
   anth_1961,anth_1962,anth_1963,anth_1964,anth_1965,
   anth_1966,anth_1967,anth_1968,anth_1969,anth_1970, 
   anth_1971,anth_1972,anth_1973,anth_1974,anth_1975,
   anth_1976,anth_1977,anth_1978,anth_1979,anth_1980, 
   anth_1981,anth_1982,anth_1983,anth_1984,anth_1985,
   anth_1986,anth_1987,anth_1988,anth_1989,anth_1990, 
   anth_1991,anth_1992,anth_1993,anth_1994,anth_1995,
   anth_1996,anth_1997,anth_1998,anth_1999,anth_2000, 
   anth_2001,anth_2002,anth_2003,anth_2004,anth_2005,
   anth_2006,anth_2007,anth_2008,anth_2009,anth_2010, 
   anth_2011,anth_2012,anth_2013,anth_2014,anth_2015,
   anth_2016,anth_2017,anth_2018,anth_2019,anth_2020, 
   anth_2021,anth_2022,anth_2023,g_area)
```



```{r}
anthromes_area_summary <- as.data.frame(cbind(
  ### 1900
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1900"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1900"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1900"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wildlands
  ), 
  ### 1910
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1910"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1910"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1910"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wildlands
  ), 
### 1920
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1920"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1920"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1920"] %in% c("61","62","63"),"area"])  ## classification 53 is remote woodlands/forests
  ), 
  ### 1930
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1930"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1930"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1930"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 1940
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1940"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1940"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1940"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1950
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1950"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1950"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1950"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 1951
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1951"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1951"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1951"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1952
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1952"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1952"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1952"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 1953
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1953"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1953"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1953"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1954
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1954"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1954"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1954"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 1955
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1955"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1955"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1955"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1956
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1956"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1956"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1956"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 

### 1957
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1957"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1957"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1957"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1958
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1958"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1958"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1958"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wildlands
  ), 
### 1959
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1959"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1959"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1959"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1960
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1960"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1960"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1960"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 

### 1961
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1961"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1961"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1961"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1962
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1962"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1962"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1962"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 1963
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1963"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1963"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1963"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1964
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1964"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1964"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1964"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 1965
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1965"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1965"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1965"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1966
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1966"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1966"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1966"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 

### 1967
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1967"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1967"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1967"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1968
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1968"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1968"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1968"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 1969
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1969"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1969"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1969"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1970
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1970"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1970"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1970"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 1971
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1971"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1971"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1971"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1972
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1972"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1972"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1972"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 1973
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1973"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1973"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1973"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1974
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1974"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1974"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1974"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 1975
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1975"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1975"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1975"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1976
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1976"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1976"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1976"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 

### 1977
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1977"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1977"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1977"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1978
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1978"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1978"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1978"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 1979
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1979"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1979"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1979"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1980
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1980"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1980"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1980"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 1981
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1981"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1981"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1981"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1982
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1982"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1982"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1982"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 1983
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1983"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1983"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1983"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1984
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1984"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1984"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1984"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 1985
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1985"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1985"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1985"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1986
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1986"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1986"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1986"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 

### 1987
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1987"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1987"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1987"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1988
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1988"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1988"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1988"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 1989
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1989"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1989"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1989"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1990
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1990"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1990"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1990"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 

### 1991
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1991"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1991"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1991"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1992
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1992"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1992"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1992"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 1993
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1993"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1993"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1993"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1994
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1994"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1994"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1994"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 1995
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1995"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1995"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1995"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1996
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1996"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1996"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1996"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 

### 1997
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1997"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1997"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1997"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 1998
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1998"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1998"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1998"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 1999
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"1999"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"1999"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"1999"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 2000
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2000"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2000"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2000"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 

### 2001
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2001"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2001"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2001"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 2002
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2002"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2002"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2002"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 2003
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2003"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2003"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2003"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 2004
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2004"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2004"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2004"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 2005
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2005"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2005"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2005"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 2006
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2006"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2006"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2006"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 

### 2007
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2007"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2007"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2007"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 2008
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2008"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2008"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2008"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 2009
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2009"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2009"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2009"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 2010
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2010"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2010"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2010"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 


### 2011
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2011"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2011"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2011"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 2012
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2012"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2012"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2012"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 2013
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2013"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2013"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2013"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 2014
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2014"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2014"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2014"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 2015
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2015"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2015"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2015"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 2016
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2016"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2016"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2016"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 

### 2017
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2017"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2017"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2017"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 2018
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2018"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2018"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2018"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
### 2019
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2019"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2019"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2019"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ), 
  ### 2020
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2020"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2020"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2020"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ),
### 2021
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2021"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2021"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2021"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ),
  ### 2022
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2022"] %in% c("11","12", "21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2022"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2022"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  ),
  ### 2023
  c(
    sum(HYDE_anthromes[HYDE_anthromes[,"2023"] %in% c("11","12","21","22","23","24","31","32","33","34","41","42","43"),"area"]),  ## classifications 11-43 are intensive Anthromes
    sum(HYDE_anthromes[HYDE_anthromes[,"2023"] %in% c("51","52","53","54"),"area"]),  ## classification 51-54 are cultured Anthromes 
    sum(HYDE_anthromes[HYDE_anthromes[,"2023"] %in% c("61","62","63"),"area"])  ## classifications 61-63 are wild lands
  )
))

### colnames and rownames 
colnames(anthromes_area_summary) <- c(seq(1900, 1950, by = 10), seq(1951, 2023, by = 1))
rownames(anthromes_area_summary) <- c("intensive","cultured","wild")

### add class variable -- used in melt later on
anthromes_area_summary$class <- as.factor(c("intensive","cultured","wild"))
```

```{r}
### make long
anthromes_summary_long <- reshape2::melt(anthromes_area_summary[1:11,], id.vars = "class", variable.name = "year", value.name = "area")

anthromes_summary_long$class <- factor(anthromes_summary_long$class, levels = rev(c("intensive","cultured","wild")))

anthromes_summary_long$area <- as.numeric(anthromes_summary_long$area)

```


```{r}
## subset dataset 
ds_4plotting <- anthromes_summary_long[anthromes_summary_long$class %in% c("intensive","cultured","wild"),]
ds_4plotting$area <- ds_4plotting$area/1000000 ### diivet to get ha *10^6
ds_4plotting$year <- as.numeric(levels(ds_4plotting$year))[ds_4plotting$year] 
```

```{r}
gg.anthrome_area2 <-  ggplot(aes(x = year, y = area, color = class, group = class, shape = class), data = ds_4plotting) + 
                      geom_hline(yintercept = 0, lty = 3, col = "gray") +  
                      geom_hline(yintercept = 20, lty = 3, col = "gray") +                    
                      geom_hline(yintercept = 40, lty = 3, col = "gray") +
                      geom_hline(yintercept = 60, lty = 3, col = "gray") +
                      geom_point(size = 1) +
                      geom_line() + 
                      theme_classic() + 
                      scale_color_manual(values = rev(c("#efcb68", "#70a0af", "#160c28")), 
                                         name = "3-class Anthrome") + 
                      scale_shape_manual(values = c(15, 16, 17), name = "3-class Anthrome") +
                      ylab(expression("area (ha  10"^6*")")) + 
                      theme(legend.position = "top") #+ ylim(c(0,18))

# tiff(file = "anthrome_areas_2.16.2024.tiff", width = 25, height = 10, units = "cm", compression = "lzw", res = 600)
gg.anthrome_area2
# dev.off()
```

# Fig 2

```{r}
library(readr)

## read in regionalized population estimates from HYDE 3.2
popc_r <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/baseline/txt/popc_r.txt")
urbc_r <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/baseline/txt/urbc_r.txt")

## pop density
popd_r <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/baseline/txt/popd_r.txt")
## same for urban population estimates
uopp_r <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/baseline/txt/uopp_r.txt")
## rangeland
conv_rangeland_r <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/baseline/txt/conv_rangeland_r.txt")
## cropland
cropland_r <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/baseline/txt/cropland_r.txt")
## grazingland
grazing_r <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/baseline/txt/grazing_r.txt")
## irrigated lands - no rice
ir_norice_r <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/baseline/txt/ir_norice_r.txt")
## irrigated lands with Rice
ir_rice_r <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/baseline/txt/ir_rice_r.txt")

### pasturelands
pasture_r <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/baseline/txt/pasture_r.txt")
## rangelands
rangeland_r <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/baseline/txt/rangeland_r.txt")
## rainfed rice
rf_rice_r <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/baseline/txt/rf_rice_r.txt")
## rainfed_norice
rf_norice_r <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/baseline/txt/rf_norice_r.txt")
## rural poplulation
rurc_r  <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/baseline/txt/rurc_r.txt")
## total irrigated area
tot_irri_r  <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/baseline/txt/tot_irri_r.txt")
## total rainfed area
tot_rainfed_r  <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/baseline/txt/tot_rainfed_r.txt")
## total rice area
tot_rice_r  <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/baseline/txt/tot_rice_r.txt")


#######################################################################################################################################################
####### READ LOWER -- this is for the CONFIDENCE INTERVAL stuff 
#######################################################################################################################################################

popc_r_lower <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/lower/popc_r.txt")
urbc_r_lower <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/lower/urbc_r.txt")
## pop density
popd_r_lower <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/lower/popd_r.txt")
## same for urban population estimates
uopp_r_lower <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/lower/uopp_r.txt")
## rangeland
conv_rangeland_r_lower <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/lower/conv_rangeland_r.txt")
## cropland
cropland_r_lower <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/lower/cropland_r.txt")
## grazingland
grazing_r_lower <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/lower/grazing_r.txt")
## irrigated lands - no rice
ir_norice_r_lower <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/lower/ir_norice_r.txt")
## irrigated lands with Rice
ir_rice_r_lower <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/lower/ir_rice_r.txt")

### pasturelands
pasture_r_lower <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/lower/pasture_r.txt")
## rangelands
rangeland_r_lower <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/lower/rangeland_r.txt")
## rainfed rice
rf_rice_r_lower <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/lower/rf_rice_r.txt")
## rainfed_norice
rf_norice_r_lower <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/lower/rf_norice_r.txt")
## rural poplulation
rurc_r_lower  <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/lower/rurc_r.txt")
## total irrigated area
tot_irri_r_lower  <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/lower/tot_irri_r.txt")
## total rainfed area
tot_rainfed_r_lower  <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/lower/tot_rainfed_r.txt")
## total rice area
tot_rice_r_lower  <- read_table("C:/Users/JamesHogan/Box/James.Hogan_Workspace/ANTHROMES/HYDE_data_from_Kees_Goldewijk/lower/tot_rice_r.txt")
```

```{r}
#############################################################################################################################################
##########


### TWO CLASSIFICATIONS:
# Temperate vs. tropical -- based on IMAGE 3.0 regions, which are tropical vs. extratropical (i.e., temperate)?  see:  https://models.pbl.nl/image/index.php/IMAGE_framework/IMAGE_3.0_in_a_nutshell
temperate <- c(1,2,11,12,13,14,15,16,19,20,23,27,28)
tropical <- c(3,4,5,6,7,8,9,10,17,18,21,22,24,25,26)


years <- c(seq(1900,1950, by = 10), seq(1951,2023, by = 1))
```

## population 
### compute summaries
```{r}
#### POPULATION SUMMARY
HYDE_3.3_popSumm  <- data.frame("year" = years,
                                "land_use" = rep(c(rep("total", length(years)), rep("urban", length(years)), rep("rural", length(years))), 6),
                                #### POPULATION
                                "global" = c(colSums(popc_r[!popc_r$region == "Total", as.character(years)]), 
                                             colSums(urbc_r[!urbc_r$region == "Total", as.character(years)]), 
                                             colSums(rurc_r[!rurc_r$region == "Total", as.character(years)])),
                                "global.lower" = c(colSums(popc_r_lower[!popc_r$region == "Total", as.character(years)]), 
                                                   colSums(urbc_r_lower[!urbc_r$region == "Total", as.character(years)]), 
                                                   colSums(rurc_r_lower[!rurc_r$region == "Total", as.character(years)])),
                                "tropical" = c(colSums(popc_r[popc_r$region %in% tropical, as.character(years)]), 
                                               colSums(urbc_r[urbc_r$region %in% tropical, as.character(years)]), 
                                               colSums(rurc_r[rurc_r$region %in% tropical, as.character(years)])),
                                "tropical.lower" = c(colSums(popc_r_lower[popc_r_lower$region %in% tropical, as.character(years)]), 
                                                     colSums(urbc_r_lower[urbc_r_lower$region %in% tropical, as.character(years)]), 
                                                     colSums(rurc_r_lower[rurc_r_lower$region %in% tropical, as.character(years)])),
                                "temperate" = c(colSums(popc_r[popc_r$region %in% temperate, as.character(years)]),
                                                colSums(urbc_r[urbc_r$region %in% temperate, as.character(years)]),  
                                                colSums(rurc_r[rurc_r$region %in% temperate, as.character(years)])), 
                                "temperate.lower" = c(colSums(popc_r_lower[popc_r$region %in% temperate, as.character(years)]),
                                                      colSums(urbc_r_lower[urbc_r$region %in% temperate, as.character(years)]),  
                                                      colSums(rurc_r_lower[rurc_r$region %in% temperate, as.character(years)])))
                          
## reshape to long 
popSumm_long <- reshape2::melt(HYDE_3.3_popSumm[,c("year", "land_use", "global", "tropical", "temperate")], id.vars = c("land_use","year"))
popSumm_long$land_use <- factor(popSumm_long$land_use, levels = c("total", "urban", "rural"))
        
### for lower
popSumm_long_lower <- reshape2::melt(HYDE_3.3_popSumm[,c("year", "land_use", "global.lower", "tropical.lower", "temperate.lower")], id.vars = c("land_use","year"))
popSumm_long_lower$land_use <- factor(popSumm_long_lower$land_use, levels = c("total", "urban", "rural"))


### add lower to popSumm_long
popSumm_long$lower <-  popSumm_long_lower$value
popSumm_long$diff <-  popSumm_long$value - popSumm_long$lower
```

### plotting 
```{r}

###############
library(ggplot2)                                

gg_pop <- ggplot(aes(x = year, y = value), data = popSumm_long) + 
          geom_hline(yintercept = 0, lty = 3) +
          geom_hline(yintercept = 2e9, lty = 3) +
          geom_hline(yintercept = 4e9, lty = 3) +
          geom_hline(yintercept = 6e9, lty = 3) +
          geom_hline(yintercept = 8e9, lty = 3) +
          geom_ribbon(aes(ymin = value + diff, ymax = value - diff, fill = variable), alpha = 0.5, color = NA) +
          geom_line(aes(color = variable)) + theme_classic() +
          facet_wrap(vars(land_use)) + ylab("population") + theme(legend.position = c(0.9,0.75)) +
          scale_color_manual(values = c("#4c5b5c", "#3772ff", "#df2935"), name = NULL) +
          scale_fill_manual(values = c("#4c5b5c", "#3772ff", "#df2935"), name = NULL) +
          scale_y_continuous(labels = label_number(scale_cut = cut_short_scale(space = TRUE)))
            
gg_pop
```

## Land Cover 1
### compute summaries 
```{r}
## land cover SUMMARY

## classificaiton #1 -- pasture vs. grazing vs. rangeland vs. cropland 

HYDE_3.3_land_Summary <- data.frame("year" = years,
                                    "land_type" = rep(c(rep("cropland", length(years)), #rep("grazing land", length(years)), 
                                                        rep("rainfed area", length(years)), rep("rangeland", length(years))), 6),
                                    "global" = c(unlist(cropland_r[cropland_r$region == "Total",as.character(years)]), 
                                                 #colSums(grazing_r[as.character(years)]),
                                                 unlist(tot_rainfed_r[tot_rainfed_r$region == "Total",as.character(years)]), 
                                                 unlist(rangeland_r[rangeland_r$region == "Total",as.character(years)])),
                                    "global.lower" = c(unlist(cropland_r_lower[cropland_r_lower$region == "Total",as.character(years)]), 
                                                       #colSums(grazing_r[as.character(years)]),
                                                       unlist(tot_rainfed_r_lower[tot_rainfed_r_lower$region == "Total",as.character(years)]), 
                                                       unlist(rangeland_r_lower[rangeland_r_lower$region == "Total",as.character(years)])),
                                    "tropical" = c(colSums(cropland_r[cropland_r$region %in% tropical, as.character(years)]), 
                                                   #colSums(grazing_r[grazing_r$region %in% tropical, as.character(years)]), 
                                                   colSums(tot_rainfed_r[tot_rainfed_r$region %in% tropical, as.character(years)]), 
                                                   colSums(rangeland_r[rangeland_r$region %in% tropical, as.character(years)])),
                                    "tropical.lower" = c(colSums(cropland_r_lower[cropland_r_lower$region %in% tropical, as.character(years)]), 
                                                         #colSums(grazing_r[grazing_r$region %in% tropical, as.character(years)]), 
                                                         colSums(tot_rainfed_r_lower[tot_rainfed_r_lower$region %in% tropical, as.character(years)]), 
                                                         colSums(rangeland_r_lower[rangeland_r_lower$region %in% tropical, as.character(years)])),
                                    "temperate" = c(colSums(cropland_r[cropland_r$region %in% temperate, as.character(years)]),
                                                    #colSums(grazing_r[grazing_r$region %in% temperate, as.character(years)]),  
                                                    colSums(tot_rainfed_r[tot_rainfed_r$region %in% temperate, as.character(years)]),
                                                    colSums(rangeland_r[rangeland_r$region %in% temperate, as.character(years)])), 
                                    "temperate.lower" = c(colSums(cropland_r_lower[cropland_r_lower$region %in% temperate, as.character(years)]),
                                                         #colSums(grazing_r[grazing_r$region %in% temperate, as.character(years)]),  
                                                         colSums(tot_rainfed_r_lower[tot_rainfed_r_lower$region %in% temperate, as.character(years)]),
                                                         colSums(rangeland_r_lower[rangeland_r_lower$region %in% temperate, as.character(years)])))


landSumm_long <- reshape2::melt(HYDE_3.3_land_Summary[,c("year", "land_type", "global", "tropical", "temperate")], 
                                id.vars = c("land_type", "year"))
landSumm_long$land_type <- factor(landSumm_long$land_type, levels = c("cropland", "rainfed area", "rangeland"))

## reshape to long 
landSumm_long_lower <- reshape2::melt(HYDE_3.3_land_Summary[,c("year", "land_type", "global.lower", "tropical.lower", "temperate.lower")], 
                                      id.vars = c("land_type","year"))
landSumm_long_lower$land_type <- factor(HYDE_3.3_land_Summary$land_type, levels = c("cropland", "rainfed area", "rangeland"))


### add lower to popSumm_long
landSumm_long$lower <-  landSumm_long_lower$value
landSumm_long$diff <-  landSumm_long$value - landSumm_long$lower
```

### plotting 
```{r}
library(scales)

gg_land <- ggplot(aes(x = year, y = value), data = landSumm_long) + 
           geom_hline(yintercept = 0, lty = 3) +
           geom_hline(yintercept = 5e6, lty = 3) +
           geom_hline(yintercept = 10e6, lty = 3) +
           geom_hline(yintercept = 15e6, lty = 3) +
           geom_hline(yintercept = 20e6, lty = 3) +
           geom_ribbon(aes(ymin = value + diff, ymax = value - diff, fill = variable), alpha = 0.5, color = NA) +
           geom_line(aes(color = variable)) + theme_classic() +
           facet_wrap(vars(land_type), nrow = 1) + ylab("land area (ha)") + 
           theme(legend.position = c(0.6,0.7),
                 axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
           scale_color_manual(values = c("#4c5b5c", "#3772ff", "#df2935"), name = NULL)  +
           scale_fill_manual(values = c("#4c5b5c", "#3772ff", "#df2935"), name = NULL) +
           scale_y_continuous(labels = label_number(scale_cut = cut_short_scale(space = TRUE)))

# tiff(filename = "HYDE3.3_LandClass2.tiff", width = 18, height = 6, compression = "lzw", res = 600, units = "cm")
gg_land
# dev.off()
```

## Land Cover 2
### compute summaries 

```{r}
## classificaiton #2 -- built up area, irrigated lands, rainfed lands
HYDE_3.3_land_SummaryB <- data.frame("year" = years,
                                     "land_type" = rep(c(rep("built up area", length(years)), rep("irrigated area", length(years)), 
                                                         rep("pastureland", length(years))), 3),
                                     "global" = c(unlist(uopp_r[uopp_r$region == "Total",as.character(years)]), 
                                                  unlist(tot_irri_r[tot_irri_r$region == "Total",as.character(years)]), 
                                                  unlist(pasture_r[pasture_r$region == "Total",as.character(years)])),
                                     "global.lower" = c(unlist(uopp_r_lower[uopp_r_lower$region == "Total",as.character(years)]), 
                                                        unlist(tot_irri_r_lower[tot_irri_r_lower$region == "Total",as.character(years)]), 
                                                        unlist(pasture_r_lower[pasture_r_lower$region == "Total",as.character(years)])),
                                     "tropical" = c(colSums(uopp_r[uopp_r$region %in% tropical, as.character(years)]), 
                                                    colSums(tot_irri_r[tot_irri_r$region %in% tropical, as.character(years)]), 
                                                    colSums(pasture_r[pasture_r$region %in% tropical, as.character(years)])),
                                     "tropical.lower" = c(colSums(uopp_r_lower[uopp_r_lower$region %in% tropical, as.character(years)]), 
                                                          colSums(tot_irri_r_lower[tot_irri_r_lower$region %in% tropical, as.character(years)]), 
                                                          colSums(pasture_r_lower[pasture_r_lower$region %in% tropical, as.character(years)])),
                                     "temperate" = c(colSums(uopp_r[uopp_r$region %in% temperate, as.character(years)]),
                                                     colSums(tot_irri_r[tot_irri_r$region %in% temperate, as.character(years)]),  
                                                     colSums(pasture_r[pasture_r$region %in% temperate, as.character(years)])), 
                                     "temperate.lower" = c(colSums(uopp_r_lower[uopp_r_lower$region %in% temperate, as.character(years)]),
                                                           colSums(tot_irri_r_lower[tot_irri_r_lower$region %in% temperate, as.character(years)]),  
                                                           colSums(pasture_r_lower[pasture_r_lower$region %in% temperate, as.character(years)])))


landSumm_long2B <- reshape2::melt(HYDE_3.3_land_SummaryB[,c("year", "land_type", "global", "tropical", "temperate")], 
                                  id.vars = c("land_type", "year"))
landSumm_long2B$land_type <- factor(landSumm_long2B$land_type , levels = c("built up area", "irrigated area", "pastureland"))


## reshape to long 
landSumm_long2B_lower <- reshape2::melt(HYDE_3.3_land_SummaryB[,c("year", "land_type", "global.lower", "tropical.lower", "temperate.lower")], 
                                        id.vars = c("land_type","year"))
landSumm_long2B_lower$land_type <- factor(HYDE_3.3_land_SummaryB$land_type, levels = c("cropland", "rainfed area", "rangeland"))


### add lower to landSumm_long
landSumm_long2B$lower <- landSumm_long2B_lower$value
landSumm_long2B$diff <- landSumm_long2B$value - landSumm_long2B$lower
```

### plotting
```{r}
gg_landB <- ggplot(aes(x = year, y = value), data = landSumm_long2B) + 
            geom_hline(yintercept = 0, lty = 3) +
            geom_hline(yintercept = 2e6, lty = 3) +
            geom_hline(yintercept = 4e6, lty = 3) +
            geom_hline(yintercept = 6e6, lty = 3) +
            geom_hline(yintercept = 8e6, lty = 3) + 
            geom_ribbon(aes(ymin = value + diff, ymax = value - diff, fill = variable), alpha = 0.5, color = NA) +
            geom_line(aes(color = variable)) + theme_classic() +
            facet_wrap(vars(land_type)) + ylab("land area (ha)") + 
            theme(legend.position = c(0.1,0.7),
                  axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
            scale_color_manual(values = c("#4c5b5c", "#3772ff", "#df2935"), name = NULL) +
            scale_fill_manual(values = c("#4c5b5c", "#3772ff", "#df2935"), name = NULL) +
            scale_y_continuous(labels = label_number(scale_cut = cut_short_scale(space = TRUE)))


# tiff(filename = "HYDE3.3_LandClass3_sameScale.tiff", width = 18, height = 6, compression = "lzw", res = 600, units = "cm")
gg_landB
# dev.off()


## classificaiton #3 -- built up area, irrigated lands (SPLITTING IRRIGATED LANDS INTO RICE AND OTHERS), rainfed lands
HYDE_3.3_land_SummaryC <- data.frame("year" = years,
                                     "land_type" = rep(c(rep("built up area", length(years)), rep("total rice area", length(years)), 
                                                         rep("rainfed - no rice", length(years)), rep("rainfed - rice", length(years)), 
                                                         rep("irrigated - no rice", length(years)), rep("irrigated - rice", length(years))), 6),
                                     "global" = c(colSums(uopp_r[as.character(years)]), 
                                                  colSums(tot_rice_r[as.character(years)]), 
                                                  colSums(rf_norice_r[as.character(years)]),
                                                  colSums(rf_rice_r[as.character(years)]),
                                                  colSums(ir_norice_r[as.character(years)]),
                                                  colSums(ir_rice_r[as.character(years)])),
                                     "global.lower" = c(colSums(uopp_r_lower[as.character(years)]), 
                                                        colSums(tot_rice_r_lower[as.character(years)]), 
                                                        colSums(rf_norice_r_lower[as.character(years)]),
                                                        colSums(rf_rice_r_lower[as.character(years)]),
                                                        colSums(ir_norice_r_lower[as.character(years)]),
                                                        colSums(ir_rice_r_lower[as.character(years)])),
                                    "tropical" = c(colSums(uopp_r[uopp_r$region %in% tropical, as.character(years)]), 
                                                    colSums(tot_rice_r[tot_rice_r$region %in% tropical, as.character(years)]), 
                                                    colSums(rf_norice_r[rf_norice_r$region %in% tropical, as.character(years)]), 
                                                    colSums(rf_rice_r[rf_rice_r$region %in% tropical, as.character(years)]), 
                                                    colSums(ir_norice_r[ir_norice_r$region %in% tropical, as.character(years)]), 
                                                    colSums(ir_rice_r[ir_rice_r$region %in% tropical, as.character(years)])),
                                    "tropical.lower" = c(colSums(uopp_r_lower[uopp_r_lower$region %in% tropical, as.character(years)]), 
                                                         colSums(tot_rice_r_lower[tot_rice_r_lower$region %in% tropical, as.character(years)]), 
                                                         colSums(rf_norice_r_lower[rf_norice_r_lower$region %in% tropical, as.character(years)]), 
                                                         colSums(rf_rice_r_lower[rf_rice_r_lower$region %in% tropical, as.character(years)]), 
                                                         colSums(ir_norice_r_lower[ir_norice_r_lower$region %in% tropical, as.character(years)]), 
                                                         colSums(ir_rice_r_lower[ir_rice_r_lower$region %in% tropical, as.character(years)])),
                                     "temperate" = c(colSums(uopp_r[uopp_r$region %in% temperate, as.character(years)]),
                                                     colSums(tot_rice_r[tot_rice_r$region %in% temperate, as.character(years)]), 
                                                     colSums(rf_norice_r[rf_norice_r$region %in% temperate, as.character(years)]), 
                                                     colSums(rf_rice_r[rf_rice_r$region %in% temperate, as.character(years)]), 
                                                     colSums(ir_norice_r[ir_norice_r$region %in% temperate, as.character(years)]), 
                                                     colSums(ir_rice_r[ir_rice_r$region %in% temperate, as.character(years)])), 
                                    "temperate.lower" = c(colSums(uopp_r_lower[uopp_r_lower$region %in% temperate, as.character(years)]),
                                                          colSums(tot_rice_r_lower[tot_rice_r_lower$region %in% temperate, as.character(years)]), 
                                                          colSums(rf_norice_r_lower[rf_norice_r_lower$region %in% temperate, as.character(years)]), 
                                                          colSums(rf_rice_r_lower[rf_rice_r_lower$region %in% temperate, as.character(years)]), 
                                                          colSums(ir_norice_r_lower[ir_norice_r_lower$region %in% temperate, as.character(years)]), 
                                                          colSums(ir_rice_r_lower[ir_rice_r_lower$region %in% temperate, as.character(years)])))


landSumm_long2C <- reshape2::melt(HYDE_3.3_land_SummaryC[,c("year", "land_type", "global", "tropical", "temperate")], id.vars = c("land_type", "year"))
landSumm_long2C$land_type <- factor(landSumm_long2C$land_type , levels = c("built up area", "total rice area", "rainfed - no rice", 
                                                                           "rainfed - rice", "irrigated - rice", "irrigated - no rice"))


landSumm_long2C_lower <- reshape2::melt(HYDE_3.3_land_SummaryC[,c("year", "land_type", "global.lower", "tropical.lower", "temperate.lower")], 
                                        id.vars = c("land_type", "year"))
landSumm_long2C_lower$land_type <- factor(landSumm_long2C$land_type , levels = c("built up area", "total rice area", "rainfed - no rice", 
                                                                                 "rainfed - rice", "irrigated - rice", "irrigated - no rice"))


### add lower to landSumm_long2C
landSumm_long2C$lower <- landSumm_long2C_lower$value
landSumm_long2C$diff <- landSumm_long2C$value - landSumm_long2C$lower


gg_land4 <- ggplot(aes(x = year, y = value), data = landSumm_long2C) + 
            geom_ribbon(aes(ymin = value + diff, ymax = value - diff, fill = variable), alpha = 0.5, color = NA) +
            geom_line(aes(color = variable)) + theme_classic() +
            facet_wrap(vars(land_type)) + ylab("land area (ha)") + theme(legend.position = "top") +
            scale_color_manual(values = c("#4c5b5c", "#3772ff", "#df2935"), name = NULL) +
            scale_fill_manual(values = c("#4c5b5c", "#3772ff", "#df2935"), name = NULL) 

# tiff(filename = "HYDE3.3_LandClass4_sameScale.tiff", width = 18, height = 12, compression = "lzw", res = 600, units = "cm")
gg_land4
# dev.off()
```



## ALL plotting together 
```{r}
#### PUTTING TOGETHER
library(ggpubr)

# tiff(filename = "HYDE3.3_ScenarioSummary_fixedScales.tiff", width = 18, height = 18, compression = "lzw", res = 600, units = "cm")
ggarrange(gg_pop + theme(axis.text.x = element_blank(), 
                          axis.title.x = element_blank(),
                          legend.position = "none"),
          gg_landB  + theme(axis.text.x = element_blank(), 
                            axis.title.x = element_blank(),
                            legend.position = "none"), 
          gg_land + theme(legend.position = "none"),  
          nrow = 3, ncol = 1, heights = c(1,1,1.25), labels = "AUTO", common.legend = T, align = "v")
# dev.off()

```

## Land Summary XX
### computing summaries 
```{r}
#################################


HYDE_3.3_land_SummaryX <- data.frame("year" = years,
                                     "land_type" = rep(c(rep("built up area", length(years)), rep("irrigated area", length(years)), 
                                                         rep("rainfed area", length(years)), rep("cropland", length(years)),
                                                         rep("rangeland", length(years)), rep("pastureland", length(years))), 3),
                                     "global" = c(colSums(uopp_r[as.character(years)]), 
                                                  colSums(tot_irri_r[as.character(years)]),
                                                  colSums(tot_rainfed_r[as.character(years)]),
                                                  colSums(cropland_r[as.character(years)]),
                                                  colSums(rangeland_r[as.character(years)]), 
                                                  colSums(pasture_r[as.character(years)])),
                                     "tropical" = c(colSums(uopp_r[uopp_r$region %in% tropical, as.character(years)]), 
                                                    colSums(tot_irri_r[tot_irri_r$region %in% tropical, as.character(years)]), 
                                                    colSums(tot_rainfed_r[tot_rainfed_r$region %in% tropical, as.character(years)]), 
                                                    colSums(cropland_r[cropland_r$region %in% tropical, as.character(years)]), 
                                                    colSums(rangeland_r[rangeland_r$region %in% tropical, as.character(years)]), 
                                                    colSums(pasture_r[pasture_r$region %in% tropical, as.character(years)])),
                                     "temperate" = c(colSums(uopp_r[uopp_r$region %in% temperate, as.character(years)]),
                                                     colSums(tot_irri_r[tot_irri_r$region %in% temperate, as.character(years)]),  
                                                     colSums(tot_rainfed_r[tot_rainfed_r$region %in% temperate, as.character(years)]), 
                                                     colSums(cropland_r[cropland_r$region %in% temperate, as.character(years)]),
                                                     colSums(rangeland_r[rangeland_r$region %in% temperate, as.character(years)]),
                                                     colSums(pasture_r[pasture_r$region %in% temperate, as.character(years)])))


landSumm_longX <- reshape2::melt(HYDE_3.3_land_SummaryX, id.vars = c("land_type", "year"))
landSumm_longX$land_type <- factor(landSumm_longX$land_type , levels = c("built up area", "irrigated area", "rainfed area", 
                                                                         "cropland", "rangeland", "pastureland"))

```

###plotting
```{r}
gg_landX <- ggplot(aes(x = year, y = value), data = landSumm_longX) + 
  geom_line(aes(color = variable), lwd = 1) + theme_classic() +
  facet_wrap(vars(land_type)) + ylab("land area (ha)") + 
  theme(legend.position = c(0.1,0.7),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  scale_color_manual(values = c("#4c5b5c", "#3772ff", "#df2935"), name = NULL) +
  scale_y_continuous(labels = label_number(scale_cut = cut_short_scale(space = TRUE)))


gg_landX

```




